---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

```{r}
path <- getwd()
url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(url, file.path(path, "dataFiles.zip"))
unzip(zipfile = file.path(path, "dataFiles.zip"))
df<-read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
```{r,echo=TRUE}
library(ggplot2)
df$date<-as.Date(df$date)
stepd<-tapply(df$steps,df$date,sum,na.rm=T)
stepd<-as.data.frame(stepd)  
names(stepd)<-c("step")
```

```{r}
ggplot(stepd,aes(x=step))+geom_histogram()
```

```{r, echo=TRUE}
mean<-mean(stepd$step)
median<-median(stepd$step)
```
We remove NAs from the dataset
Therefore, the mean steps taken each day is `r mean`. The median steps taken each day is `r median`

## What is the average daily activity pattern?

```{r echo=TRUE}
stepi<-tapply(df$steps,df$interval,mean,na.rm=T)
stepi<-as.data.frame(stepi)   
stepi<-data.frame(interval=unique(df$interval),steps=stepi$stepi)
ggplot(stepi,aes(x=interval,y=steps))+geom_line(size=1.0)
int<-stepi[order(stepi$steps,decreasing = TRUE),][1,][1]
```
on average across all the days in the dataset, interval `r int` contains the maximum number of steps

## Imputing missing values
There is a total of `r sum(is.na(df$steps))` NAs in the dataset.  
I used the mean for that 5-minute interval to as the stratergy.  
New dataset is the newdf as below.
```{r}
u<-df[which(is.na(df$steps)),]
m<-merge(u,stepi,by="interval")
m<-data.frame(steps=m$steps.y,date=m$date,interval=m$interval)
m<-m[order(m$interval,m$date),]
newdf<-df
newdf[which(is.na(newdf$steps)),][order(u$interval,u$date),]<-m
```

#Histogram
```{r}
newstepd<-tapply(newdf$steps,newdf$date,sum,na.rm=T)
newstepd<-as.data.frame(stepd)  
names(newstepd)<-c("step")
ggplot(newstepd,aes(x=step))+geom_histogram()
```

```{r, echo=TRUE}
newmean<-mean(newstepd$step)
newmedian<-median(newstepd$step)
```
Therefore, the new mean steps taken each day is `r newmean`. The median new steps taken each day is `r newmedian`. It could been seen there is no impact  of imputing missing data on the estimates of the total daily number of steps.

## Are there differences in activity patterns between weekdays and weekends?

```{r }
newdf$days<-weekdays(newdf$date)
newdf$weekend<-"weekday"
newdf$weekend[newdf$days=="Saturday"|newdf$days=="Sunday"]<-"weekend"
newdf$weekend<-as.factor(newdf$weekend)
h<-split(newdf,newdf$weekend)
h1<-h$weekday
h2<-h$weekend

stepi1<-tapply(h1$steps,h1$interval,mean,na.rm=T)
stepi1<-as.data.frame(stepi1)   
stepi2<-tapply(h2$steps,h2$interval,mean,na.rm=T)
stepi2<-as.data.frame(stepi2)   

stepi1<-data.frame(interval=unique(h1$interval),steps=stepi1$stepi1,week="weekday")
stepi2<-data.frame(interval=unique(h2$interval),steps=stepi2$stepi2,week="weekend")
stepi<-rbind(stepi1,stepi2)


ggplot(stepi,aes(x=interval,y=steps))+geom_line(size=1.0)+facet_grid(week~.)
```

So it could be seen above, that steps in weekdays are generally more in weekends. Besides, the variation in weekdays are larger than in weekend.
